Stack Safety as a Security Property

### 2020-07-02 main.tex
Updated running example

\begin{figure}
  \begin{minipage}{\textwidth}
    \begin{center}
\MemoryLabel{10em}{1.5em}{40}
\MemoryLabel{16em}{1.5em}{60}
\MemoryLabel{22em}{1.5em}{80}
$\cdots$
\memory{4}{gray}%
\memory{5}{green}[{\makebox[0pt]{Caller}}]%
\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
\memory{8}{red}[{\makebox[0pt]{Callee}}]%
\memory{4}{gray}%
~$\cdots$\\
  \end{center}
  \end{minipage}\\
~\\
~\\
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
lw r1 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}~
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
sw r0 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}
  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
      the boxes; the program fragments need comments explaining what each
      line does.}}
  \label{fig:stackunsafety}
\end{figure}

\leo{I hate the writing here, but I wanted to make a start.}

@@

In summary, we offer the following contributions:
\begin{itemize}
\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
  intuitive concepts of local state encapsulation and
  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
  property is strong enough to capture the folk intuition, we show
  that the stack unsafe examples from the literature also fail our
  property~(\Cref{sec:disc}).
\item As evidence that the property is not too strong, we show that it is
  realizible by
  a previously proposed micropolicy enforcement
  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
    substantial body of code}. We also present a variation of the
  property + policy that is lazy, easier/more efficient to implement,
  and provably weaker (\Cref{sec:lazy}).
\item A provably stronger (inductive? step-by-step/stepwise) variant
  of this property tuned for testing/verification.  We test the whole
  thing with a cool framework. \bcp{Which section?}
\end{itemize}
Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
%
We begin with some technical preliminaries in
Section~\ref{sec:prelim}.

@@

\begin{figure}[b]
  \begin{centering}
    \begin{minipage}{.8\textwidth}
\begin{verbatim}
f = {                 g(x) = {                      h(x) = {
   int x, y;                    return (x + h(17));                  return (x + 1);
  ...                 }                             }
  y = g(0);
  ...
}
\end{verbatim}
\end{minipage}
\end{centering}
\caption{High-level Program with Simple Nested Calls \bcp{Rename parameters
    of g and h to z and w}}
\label{fig:running-program}
\end{figure}

\subsection{Running Example}

\bcp{
Comments on the example:
\begin{itemize}
\item We can simplify a bit by renaming $f$ to $\mathit{main}$ and
relocating its stack to location 1000.  And maybe move its instructions to
start at 0.  (We can comment that on a real RISC-V this is where interrupt
handlers go, but POPL readers will not care too much.)  We should streamline
the example as much as possible.
\item Do we need the ``...''s in $f$?
%
\item We should also emphasize, again, that there is no heap (if there were,
we might put it above the stack or between the instructions and the stack).
%
\item We will also need to be careful about fonts ($f$ vs {\texttt f}),
capitalization of register names (sp vs SP), etc. Maybe typeset instructions
in small caps?  Maybe typeset everything that is part of a concrete machine
in small caps??
%
\item The colors are helpful in making the big page of instruction traces easier
on the eye; maybe they would be even more helpful if
corresponding colors appeared elsewhere in the example (on the instructions
themselves, e.g.).  But we need to be careful to keep it readable -- maybe
they should be background colors, like you've done, rather than text colors.
\item The first column of numbers is redundant with the second column (PC).
Suggest deleting the first.
\item Asking readers to wade through all this is asking quite a bit!  We
need to make sure we are going to get sufficient value out of the example
later (I think we can, but we should really convince ourselves) to make it
worthwhile.
\item Use ``='' instead of ``:'' in the ``summaries'' below the pictures.
And put a bit more space between the components.
\item Create a new column that just literally pronounces each instruction in
English, so people don't have to read RISC-V opcodes.
\item As discussed, we should move this to before section 2, we should
extend it with a bit more at the end so we can talk about the lazy policy
with the same example, we should add
in some bad variants (doing it here replaces the ``Examples'' from the
Discussion section), and we should consider discussing the attacker model 
here (we need to discuss it someplace in any case!)
\end{itemize}
}

Our running example consists of three functions $f$, $g$, and $h$ that
perform a nested sequence of calls. A high level C-like view of the
excerpt program can be found in Figure~\ref{fig:running-program}.
Suppose that the starting state for our running example is right
before $f$'s call to $g$. A sketch of the memory layout at that
point can be found in the top of Figure~\ref{fig:running-trace-b}:
\begin{itemize}
\item The low part of the memory (0-1000) is dedicated to instructions.
\item The stack starts at location 1000.
\item The stack of $f$ starts at 1040.
\item The location of $x$ is 1040 and stores $42$.
\item The location of $y$ is 1044 and stores $0$.
\end{itemize}
Directly below the layout, we summarize the parts of the register file
that are relevant to the example.

The rest of Figure~\ref{fig:running-trace-b} details the execution of the
program, showing the machine state at call and return points, as well
as describing the full effect of each instruction in detail.
\leo{Unoptimized, etc.}\bcp{??}

\newcommand*{\add}{\text{add}}
\newcommand*{\addi}{\text{addi}}
\newcommand*{\sw}{\text{sw}}
\newcommand*{\lw}{\text{lw}}
\newcommand*{\jal}{\text{jal}}
\newcommand*{\jalr}{\text{jalr}}
\newcommand*{\rsp}{\mathit{sp}}
\newcommand*{\rra}{\mathit{ra}}

\newcommand*{\tracerow}[5]{#1: & #2 & #3 & #4 & \text{#5}\\}
\newcommand{\negate}{\textrm{-}}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{35.2em}{1.8em}{$\downarrow$}
\MemoryLabel{35em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
\memory{2}{green}[{\makebox[0pt]{$F_{\textit{priv}}$}}]%
\memory{8}{gray}[{\makebox[0pt]{Uninitialized}}]
\MemoryLabel{-29.6em}{0.6em}{42}
\MemoryLabel{-27.4em}{0.6em}{0}
\begin{array}{cccccc}
  PC: 100 & r_0 : 0 & \rra : ? & \rsp : 1048 & r_4 : ? & r_5 : ?
  \begin{array}{l|l|l|l|l}
    & PC & \mathit{Instruction} & \mathit{Effect} &  \\
    \tracerow{1}{100}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1056}{Allocate call frame}
    \tracerow{2}{104}{\sw ~ \rsp ~ r_0 ~ \negate 8}{[1048] \leftarrow 0}{Store 0 as argument}
    \tracerow{3}{108}{\sw ~ \rsp ~ r_0 ~ \negate 4}{[1052] \leftarrow 0}{Zero out result}
    \tracerow{4}{112}{\jal ~ \rra ~ 140}{PC \leftarrow 140; ~ \rra \leftarrow 116}{Make the call}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{41em}{1.8em}{$\downarrow$}
\MemoryLabel{40.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg + result}}]%
\memory{6}{gray}[{\makebox[0pt]{Uninitialized}}]
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\begin{array}{cccccc}
  PC: 140 & r_0 : 0 & \rra : 116 & \rsp : 1056 & r_4 : ? & r_5 : ?
  \begin{array}{l|l|l|l|l}
    \tracerow{5}{140}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1060}{Allocate for return address}
    \tracerow{6}{144}{\sw ~ \rsp ~ \rra ~ \negate 4}{[1056] \leftarrow 116}{Store return address}
    \tracerow{7}{148}{\addi ~ \rsp ~ \rsp ~ 8}{\rsp \leftarrow 1068}{Allocate for call frame}
    \tracerow{8}{152}{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow 17}{Store 17 to $r_4$}
    \tracerow{9}{156}{\sw ~ \rsp ~ r_4 ~ \negate 8}{[1060] \leftarrow 17}{Store 17 as the argument}
    \tracerow{10}{160}{\sw ~ \rsp ~ r_0 ~ \negate 4}{[1064] \leftarrow 0 }{Zero out return}
    \tracerow{11}{164}{\jal ~ \rra ~ 200}{PC \leftarrow 200; ~ \rra \leftarrow 168}{Make the call to $h$}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{49em}{1.8em}{$\downarrow$}
\MemoryLabel{48.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{3}{gray}[{\makebox[0pt]{Uninitialized}}]
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}
\MemoryLabel{-22.8em}{0.6em}{0}
\MemoryLabel{-20.6em}{0.6em}{116}
\MemoryLabel{-18.4em}{0.6em}{17}
\MemoryLabel{-16.2em}{0.6em}{0}
\begin{array}{cccccc}
  PC: 200 & r_0 : 0 & \rra : 168 & \rsp : 1068 & r_4 : 17 & r_5 : ?
\begin{array}{l|l|l|l|l}
  \tracerow{12}{200}{\addi ~ \rsp ~ \rsp ~ 4}{\rsp \leftarrow 1072}{Allocate for return address}
  \tracerow{13}{204}{\sw ~ \rsp ~ \rra ~ \negate 4}{[1068] \leftarrow 168}{Store return address}
  \tracerow{14}{208}{\lw ~ r_4 ~ \rsp ~ \negate 12}{r_4 \leftarrow 17}{Load the argument of $h$ to $r_4$}
  \tracerow{15}{212}{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{Increment $r_4$ by 1}\end{array}
\MemoryLabel{1.7em}{1.8em}{0}
\MemoryLabel{21.2em}{1.8em}{1000}
\MemoryLabel{29.7em}{1.8em}{1040}
\MemoryLabel{52em}{1.8em}{$\downarrow$}
\MemoryLabel{51.8em}{3em}{SP}
\memory{6}{gray}[{\makebox[0pt]{Instructions}}]%
\memory{2}{gray}[{\makebox[0pt]{Stack}}]%
\memory{2}{green}[{\makebox[0pt]{$F_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{arg-res}}]%
\memory{3}{red}[{\makebox[0pt]{ra-arg-res}}]%
\memory{1}{cyan}%
\memory{2}{gray}[{\makebox[0pt]{Uninitialized}}]
\MemoryLabel{-29.4em}{0.6em}{42}
\MemoryLabel{-27.2em}{0.6em}{0}
\MemoryLabel{-25.0em}{0.6em}{0}

>>>>

%\begin{figure}
%  \begin{minipage}{\textwidth}
%    \begin{center}
%\MemoryLabel{10em}{1.5em}{40}
%\MemoryLabel{16em}{1.5em}{60}
%\MemoryLabel{22em}{1.5em}{80}
%$\cdots$
%\memory{4}{gray}%
%\memory{5}{green}[{\makebox[0pt]{Caller}}]%
%\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
%\memory{8}{red}[{\makebox[0pt]{Callee}}]%
%\memory{4}{gray}%
%~$\cdots$\\
%  \end{center}
%  \end{minipage}\\
%~\\
%~\\
%  \begin{minipage}{0.45\textwidth}
%    \begin{center}
%\begin{verbatim}
%jal ...
%lw r1 r0 40
%\end{verbatim}
%    \end{center}
%  \end{minipage}~
%  \begin{minipage}{0.45\textwidth}
%    \begin{center}
%\begin{verbatim}
%jal ...
%sw r0 r0 40
%\end{verbatim}
%    \end{center}
%  \end{minipage}
%  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
%      the boxes; the program fragments need comments explaining what each
%      line does.}}
%  \label{fig:stackunsafety}
%\end{figure}
% 
%\leo{I hate the writing here, but I wanted to make a start.}

@@

In summary, we offer the following contributions:
\begin{itemize}
\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
  intuitive concepts of local state encapsulation and
  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
  property is strong enough to capture the folk intuition, we show
  that the stack unsafe examples from the literature also fail our
  property~(\Cref{sec:disc}).
\item As evidence that the property is not too strong, we show that it is
  realizible by
  a previously proposed micropolicy enforcement
  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
    substantial body of code}. We also present a variation of the
  property + policy that is lazy, easier/more efficient to implement,
  and provably weaker (\Cref{sec:lazy}).
\item A provably stronger (inductive? step-by-step/stepwise) variant
  of this property tuned for testing/verification.  We test the whole
  thing with a cool framework. \bcp{Which section?}
\end{itemize}
Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
%
We begin with a running example that we're going to keep coming back
to throughout the paper (Section~\ref{sec:running-example}) and some
technical preliminaries (Section~\ref{sec:prelim}).

\section{Running Example}
\label{sec:running-example}

To help keep things concrete throughout the more technical sections
that follow, we are first going to introduce a running example. We are
going to use it to demonstrate various potentially problematic
behaviours in this section, and then show how our properties and
enforcement mechanisms preclude them later on in the paper.

The program appears in Figure~\ref{fig:running-program}. It consists
of a {\tt main} function that performs a call to some function {\tt
  f}, which in turns makes a nested call to some function {\tt g}
before returning, all while performing some simple arithmetic
in the process.

\begin{figure}[b]
  \begin{centering}
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
main = {
  int x -42, y = 0;
  y = g(0);
  print (x + y);
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
g(z) = {
  int w = h(17);
  print z;
  return (z + w);
}
\end{verbatim}
    \end{minipage}
    \begin{minipage}{.3\textwidth}
\begin{verbatim}
h(v) = {
  return (v + 1);
}

\end{verbatim}
    \end{minipage}
\end{centering}
\caption{High-level Program with Simple Nested Calls}
\label{fig:running-program}
\end{figure}

To show a trace of the execution of this program, we are going to need
a low-level view of both the code and the machine it runs on. While
our concrete implementation (Section~\leo{cite}) operates on a full
conventional RISC-V architecture, we will make a few simplifying
assumptions here to ease the exposition.

First of all, we are going to ignore the heap. We are mainly concerned
with stack interactions so that will be the main focus. \leo{We're
  going to come back to that at the end. Will we?} We are then going
to abstract away low-level details of memory layout and present a
machine that is word-addressed, with instructions starting from
location 0, a dedicated output at location 99, and lay the stack from
location 100 upwards. \leo{Explain more? How to connect to realistic
  one? Observations? DMA? Interrupt handlers?}
%
A sketch of the memory layout at that point can be found in the top of
Figure~\ref{fig:running-trace-b}: instructions for {\tt main} begin at
0, the dedicated output is empty, while the (uninitialized) stack
begins at 100. Directly below the layout, we summarize the parts of
the register file that are relevant to the example.
%
The rest of Figure~\ref{fig:running-trace-b} details the execution of the
program, showing the machine state at call and return points, as well
as describing the full effect of each instruction in detail.
\leo{Expand upon this?}

\bcp{ Comments on the example:
\begin{itemize}
\item We can simplify a bit by renaming $f$ to $\mathit{main}$ and
relocating its stack to location 1000.  And maybe move its instructions to
start at 0.  (We can comment that on a real RISC-V this is where interrupt
handlers go, but POPL readers will not care too much.)  We should streamline
the example as much as possible.
%
\item We should also emphasize, again, that there is no heap (if there were,
we might put it above the stack or between the instructions and the stack).
%
\item We will also need to be careful about fonts ($f$ vs {\texttt f}),
capitalization of register names (sp vs SP), etc. Maybe typeset instructions
in small caps?  Maybe typeset everything that is part of a concrete machine
in small caps??
%
\item The colors are helpful in making the big page of instruction traces easier
on the eye; maybe they would be even more helpful if
corresponding colors appeared elsewhere in the example (on the instructions
themselves, e.g.).  But we need to be careful to keep it readable -- maybe
they should be background colors, like you've done, rather than text colors.
\item The first column of numbers is redundant with the second column (PC).
Suggest deleting the first.
\item Asking readers to wade through all this is asking quite a bit!  We
need to make sure we are going to get sufficient value out of the example
later (I think we can, but we should really convince ourselves) to make it
worthwhile.
\item Use ``='' instead of ``:'' in the ``summaries'' below the pictures.
And put a bit more space between the components.
\item Create a new column that just literally pronounces each instruction in
English, so people don't have to read RISC-V opcodes.
\item As discussed, we should move this to before section 2, we should
extend it with a bit more at the end so we can talk about the lazy policy
with the same example, we should add
in some bad variants (doing it here replaces the ``Examples'' from the
Discussion section), and we should consider discussing the attacker model 
here (we need to discuss it someplace in any case!)
\end{itemize}
}

\newcommand*{\add}{\textsc{add}}
\newcommand*{\addi}{\textsc{addi}}
\newcommand*{\sw}{\textsc{sw}}
\newcommand*{\lw}{\textsc{lw}}
\newcommand*{\jal}{\textsc{jal}}
\newcommand*{\jalr}{\textsc{jalr}}
\newcommand*{\rsp}{\textsc{sp}}
\newcommand*{\rra}{\textsc{ra}}
\newcommand*{\rout}{\textsc{out}}

\newcounter{pcctr}
\setcounter{pcctr}{0}
\newcommand*{\row}[4]{
  \thepcctr & \stepcounter{pcctr} #1 & #2 & #3 & \text{#4}\\
}
\newcommand*{\tracerow}[5]{
  #1 & #2 & #3 & #4 & \text{#5}\\
}
\newcommand*{\summary}[6]{
\[
\begin{array}{cccccc}
  \PCname = #1 & r_0 = #2 & \rra = #3 & \rsp = #4 & r_4 = #5 & r_5 = #6
\end{array}
\]
}    
\newcommand{\negate}{\textrm{-}}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{25em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{10}{gray}[{\makebox[0pt]{Uninitialized}}]
\summary{0}{0}{?}{100}{?}{?}
  \begin{array}{c|c|c|c|c}
    PC & \mathit{Instruction} & \mathit{Description} & \mathit{Effect} & \\
    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 102}
        {Allocate private {\tt main} data}
    \row{\addi ~ r_4 ~ r_0 ~ 42}{r_4 \leftarrow r_0 + 42}{r_4 \leftarrow 42}
        {Put 42 in $r_4$}
    \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp-2] \leftarrow r_4}{[100] \leftarrow 42}
        {Set {\tt x} to 42}
    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[101] \leftarrow 0}
        {Set {\tt y} to 0}
    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 104}
        {Allocate {\tt g}'s call frame}
    \row{\sw ~ \rsp ~ r_0 ~ \negate 2}{[\rsp-2] \leftarrow r_0}{[102] \leftarrow 0}
        {Set argument to 0}
    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp-1] \leftarrow r_0}{[103] \leftarrow 0}
        {Zero out result}
    \row{\jal ~ \rra ~ 20}{??}{\PCname \leftarrow 20; ~ \rra \leftarrow \thepcctr}
        {Make the call}
  ~\\
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{6}{gray}%[{\makebox[0pt]{Uninitialized}}]
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\MemoryLabel{-38.5em}{0.75em}{0}
\MemoryLabel{-34.5em}{0.75em}{0}
\summary{20}{0}{\thepcctr}{104}{42}{?}
%\vspace*{0.2em}
\setcounter{pcctr}{20}%
  \begin{array}{c|c|c|c|c}
    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 106}
        {Allocate local {\tt g} state}
    \row{\sw ~ \rsp ~ \rra ~ \negate 2}{[\rsp - 2] \leftarrow \rra}{[104] \leftarrow 8}
        {Store return address}
    % Zero out z?
    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\rsp \leftarrow 108}
        {Allocate for call frame}
    \row{\addi ~ r_4 ~ r_0 ~ 17}{r_4 \leftarrow r_0 + 17}{r_4 \leftarrow 17}
        {Store 17 to $r_4$}
    \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[106] \leftarrow 17}
        {Store 17 as the argument {\tt v}}
    \row{\sw ~ \rsp ~ r_0 ~ \negate 1}{[\rsp - 1] \leftarrow r_0}{[107] \leftarrow 0}
        {Zero out result}
    \row{\jal ~ \rra ~ 40}{??}{PC \leftarrow 40; ~ \rra \leftarrow \thepcctr}
        {Make the call to {\tt h}}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{61.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\MemoryLabel{-38.5em}{0.75em}{0}
\MemoryLabel{-34.5em}{0.75em}{0}
\MemoryLabel{-30.5em}{0.75em}{8}
\MemoryLabel{-26.5em}{0.75em}{}
\MemoryLabel{-22.5em}{0.75em}{17}
\MemoryLabel{-18.5em}{0.75em}{0}
\summary{40}{0}{\thepcctr}{108}{17}{?}
\setcounter{pcctr}{40}
\begin{array}{c|c|c|c|c}
  \row{\addi ~ \rsp ~ \rsp ~ 1}{\rsp \leftarrow \rsp + 1}{\rsp \leftarrow 109}
      {Allocate for return address}
  \row{\sw ~ \rsp ~ \rra ~ \negate 1}{[\rsp - 1] \leftarrow \rra}{[108] \leftarrow 27}
      {Store return address}
  \row{\lw ~ r_4 ~ \rsp ~ \negate 3}{r_4 \leftarrow [\rsp - 3]}{r_4 \leftarrow 17}
      {Load the argument {\tt v} to $r_4$}
  \row{\addi ~ r_4 ~ r_4 ~ 1}{r_4 \leftarrow r_4 + 1}{r_4 \leftarrow 18}
      {Increment $r_4$ by 1}\end{array}
~\\
~\\
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{65.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
\memory{1}{red}%[{\makebox[0pt]{{\texttt h} ra}}]
\memory{1}{gray}%[{\makebox[0pt]{Uninitialized}}]
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\MemoryLabel{-38.5em}{0.75em}{0}
\MemoryLabel{-34.5em}{0.75em}{0}
\MemoryLabel{-30.5em}{0.75em}{8}
\MemoryLabel{-26.5em}{0.75em}{}
\MemoryLabel{-22.5em}{0.75em}{17}
\MemoryLabel{-18.5em}{0.75em}{0}
\MemoryLabel{-14.5em}{0.75em}{27}
\summary{40}{0}{\thepcctr}{109}{18}{?}

\begin{figure}
\[
\begin{array}{c|c|c|c|c}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 2}{[\rsp - 2] \leftarrow r_4}{[107] \leftarrow 18}
      {Store $r_4$ as the return value}
  \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\rra \leftarrow 31}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 1}{\rsp \leftarrow \rsp - 1}{\rsp \leftarrow 108}
      {Restore the stack pointer of {\tt g}}
  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 31}
      {Return to {\tt g}}
  \end{array}
\]

% After Return from H
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{61.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{2}{yellow}[{\makebox[0pt]{${\texttt g}_{priv}$}}]%
\memory{2}{yellow}[{\makebox[0pt]{{\tt h}'s call frame }}]%
\memory{2}{gray}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-46.5em}{0.75em}{42}
\MemoryLabel{-42.5em}{0.75em}{0}
\MemoryLabel{-38.5em}{0.75em}{0}
\MemoryLabel{-34.5em}{0.75em}{0}
\MemoryLabel{-30.5em}{0.75em}{8}
\MemoryLabel{-26.5em}{0.75em}{}
\MemoryLabel{-22.5em}{0.75em}{17}
\MemoryLabel{-18.5em}{0.75em}{18}
\\
\end{center}
% Starting Register File
\summary{27}{0}{\thepcctr}{108}{18}{?}
\vspace*{0.2em}

\setcounter{pcctr}{27}
% H - Body
\vspace*{0.2em}
\[
\begin{array}{c|c|c|c|c}
  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
      {Set {\tt w} to 18}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
      {Deallocate {\tt h}'s call frame}
  \row{\lw ~ r_5 ~ \rsp ~ \negate 6}{r_5 \leftarrow [\rsp - 4]}{r_5 \leftarrow 0}
      {Load argument {\tt z} $r_5$}
  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
      {Print {\tt z}}      
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
      {Add $r_4$ and $r_5$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
      {Store $r_4$ as ther return of {\tt g}}
  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
      {Deallocate {\tt g}'s local state}
  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 8}{Return to {\tt main}}
\end{array}
\]
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{6}{gray}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{0}
\MemoryLabel{-47em}{0.75em}{42}
\MemoryLabel{-43em}{0.75em}{0}
\MemoryLabel{-39em}{0.75em}{0}
\MemoryLabel{-35em}{0.75em}{18}
\\
\end{center}
% Register File
\summary{8}{0}{\thepcctr}{104}{18}{0}
%\vspace*{0.2em}
\setcounter{pcctr}{8}
\[
\begin{array}{c|c|c|c|c}
  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
      {Set {\tt y} to the result}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
      {Deallocate {\tt g}'s call frame}
  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
      {Load {\tt x} to $r_5$}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
      {Add {\tt x} and {\tt y}}
  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
      {Print {\tt x + y}}
\end{array}
\]
\vspace*{0.2em}
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{31.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{8}{gray}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{60}
\MemoryLabel{-47em}{0.75em}{60}
\MemoryLabel{-43em}{0.75em}{42}
\\
\end{center}
% Starting Register File
\summary{\thepcctr}{0}{37}{102}{60}{42}
\caption{Detailed Execution Trace}
\label{fig:running-trace-b}
\end{figure}

@@
